### 2016-02-29 Manuscript/SW_early_life_telomeres.Rmd
More major edits

All DNA extractions that passed the above criteria were diluted to 3.3 ng $\mu$l^-1^ before telomere measurement. We measured absolute telomere quantity per diploid genome for all samples using a quantitative PCR (qPCR) assay, outlined in detail by Barrett *et al.* [-@Barrett2012], with one amendment. A change in batch of SYBR green forced us to raise the annealing temperature of the telomere reaction from 58^o^C to 61^o^C for the majority of the samples (n = `r nrow(subset(dd,Objective %in% c(7:9)))`; see also below). Post-qPCR processing of samples was carried out as in Barrett *et al.* [-@Barrett2012], with the following amendments. First, averaging of technical repeats was carried out using custom-made R scripts (available as supplementary material) instead of GenEx. Second, the change in annealing temperature resulted in consistently higher, but repeatable, CQ values for the telomere reaction, so we implemented a simple correction to ensure that telomere lengths were repeatable across annealing temperatures (supplementary material). Finally, we excluded telomere lengths greater than 30kb, which maximised repeatability of our own data [see also @Bauch2012]. Inter-plate repeatability of final telomere lengths was assessed using the R package rptR [@Schielzeth2011], and for all subsequent analyses we used mean telomere length per sample where we had repeats.

@@

#####Early-life telomere length and age
Inter-plate repeatability of telomere length, based on XX samples measured at least twice, was 0.78 (CI = 0.73-0.82). Telomere length dereased with age within the first year of life (R^2^ = `r round2(summary(lm(LogTL~Agemonths,data=juv))$r.squared)`; F = `r round2(anova(lm(LogTL~Agemonths,data=juv))[1,4])`; P = `r round3(anova(lm(LogTL~Agemonths,data=juv))[1,5])`; Fig. 1A). Similarly, telomere shortening based on longitudinal data was significantly greater than zero in nestlings (one sample t-test; P < 0.001), but not in any other age group (all P > 0.5; Fig. 1B).

#####Cohort-level variation in early-life telomere dynamics
Early-life telomere length varied significantly among breeding seasons (one-way ANOVA, F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = juv))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = juv))[1,5])`). Separating temporal variaiton in telomere length among age classes revealed no significant cohort effects in chicks (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 1)))[1,4])`; P = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 1)))[1,5])`; Fig. 2A), but significant effects in fledglings (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 6)))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 6)))[1,5])`; Fig. 2B) and subadults (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 9)))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 9)))[1,5])`; Fig. 2C). Variation in median telomere length over breeding seasons in fledglings and subadults was not related temporal variation in  territory quality (linear regression, R^2^ = `r round2(summary(lm(TL~TQ,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~TQ,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~TQ,data = juvseason,na.action=na.exclude))[1,5])`), island-wide food availability (R^2^ = `r round2(summary(lm(TL~Insect,data=subset(juvseason,Insect<9),na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,5])`) or population density (R^2^ = `r round2(summary(lm(TL~Density,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~Density,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~Density,data = juvseason,na.action=na.exclude))[1,5])`).

#####Individual-level variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class and the number of non-helping subordinates in the natal territory (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lmer(LogTL~1+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE))-AICc(lmer(LogTL~Fledged+NonHelper+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE)))`). The age class effect was very strong and significant, and was present in every model in the top model set (Fig. 3A). The effect of non-helpers, on the other had did not have higher importance across the top model set than the other variables, and confidence intervals overlapped zero (Fig. 3A). Furthermore the model including non-helpers and age class model was not a significantly better fit than a model including age class alone ($\Delta$AICc = 0.06; Table S1).

The top model explaining telomere loss contained age class and tarsus length, and agian this was a significantly better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(TROC~1+(1|FieldPeriodID),data=Loss,REML = FALSE))-AICc(lmer(TROC~Fledged+Tarsus+(1|FieldPeriodID),data=Loss,REML = FALSE)))`). Fledglings lost telomeres at a lower rate compared to nestlings, while birds with longer tarsi higher rates of telomere shortening (Fig. 3B).

@@

Telomere length in early life did not affect survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~LogTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~LogTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and there was no interaction effect between telomere length and juvenile age on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~LogTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~LogTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`). To separate out cohort-level telomere length affected survival we calculated cohort-level telomere length (i.e. the mean for each cohort) and added this term to a logistic regression. However, we found no effect of telomere length on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and no interaction with juvenile age (estimate = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`)

Cox regression also showed no effect of individual-level or cohort-level telomere length on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~LogTL+Fledged)))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~LogTL)))$coefficients[1,5])`;  cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,6])`).

@@

###Discussion
Here we use the long-term study of a closed population of Seychelles warblers to assess the relationships between early-life conditions, telomere length and survival. We find that while telomere length varies

The clearest result from our study is that telomere length varies among cohorts. Very few studies have shown that temporal variation in telomere dynamics occurs in natural populations, and to our knowledge the studies that have done so were limited to just two seasons [@Mizutani2013; @Watson2015]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics does occur at the population level over longer time periods.

@@

In the Seychelles warbler we found only weak evidence that any of the social or environmental vairables we measured affect juvenile telomere length. This is surprising a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival. Possible explanations for this finding include i) temporal variation in telomere length in our data is so strong that we are unable to detect spatial trends, ii) within-cohort variation in telomere length is explained by a variable that we have not included in our analyses, or iii) that effects are generally weak and levels of noise in our telomere measurments and/or ecological data preclude detection of significant effects. Future research should therefore examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence in natural populations [@Becker2015; @Asghar2014].

Our data suggest that, in addition to age class, tarsus length and season had weak effects on juvenile telomere length (Figs 3, 4). The tarsus effect most likely reflects the fact that in passerine birds tarsus length is correlated with age during the nestling stage [@Ricklefs1976]. Telomere loss is most rapid early in life due to ongoing cell replication [@Frenck1998], and a negative correlation between telomere length and body size in early life is therefore expected. Indeed, that age class affects both tarsus length and telomere lengths can be clearly seen in our data (Fig. 4B). The difference in telomere length between birds born in summer and winter seasons is more surprising, especially given the directionality of the effect. We find that birds born in winter seasons having longer telomeres (Fig. 4A), while the opp. One possible explanation for this is that parents in good condition are more likely to breed

>>>>

All DNA extractions that passed the above criteria were diluted to 3.3 ng $\mu$l^-1^ before telomere measurement. We measured absolute telomere quantity per diploid genome for all samples using a quantitative PCR (qPCR) assay, outlined in detail by Barrett *et al.* [-@Barrett2012], with one amendment. A change in batch of SYBR green forced us to raise the annealing temperature of the telomere reaction from 58^o^C to 61^o^C for the majority of the samples (n = `r nrow(dd)`; see also below). Post-qPCR processing of samples was carried out as in Barrett *et al.* [-@Barrett2012], with the following amendments. First, averaging of technical repeats was carried out using custom-made R scripts (available as supplementary material) instead of GenEx. Second, the change in annealing temperature resulted in consistently higher, but repeatable, CQ values for the telomere reaction, so we implemented a simple correction to ensure that telomere lengths were repeatable across annealing temperatures (supplementary material). Finally, we excluded telomere lengths greater than 30kb, which maximised repeatability of our own data [see also @Bauch2012]. Inter-plate repeatability of final telomere lengths was assessed using the R package rptR [@Schielzeth2011], and for all subsequent analyses we used mean telomere length per sample where we had repeats.

@@

#####Early-life telomere dynamics
Inter-plate repeatability of telomere length, based on XX samples measured at least twice, was 0.78 (CI = 0.73-0.82). Telomere length dereased with age within the first year of life (R^2^ = `r round2(summary(lm(RTL~Agemonths,data=juv))$r.squared)`; F = `r round2(anova(lm(RTL~Agemonths,data=juv))[1,4])`; P `r round3(anova(lm(RTL~Agemonths,data=juv))[1,5])`; Fig. 1A). Longitudinal data showed that the extent of telomere loss in early life was also greatest in the nestling phase, and decreased with age (R^2^ = `r round2(summary(lm(TROC~Agemonths,data=Loss))$r.squared)`; F = `r round2(anova(lm(TROC~Agemonths,data=Loss))[1,4])`; P `r round3(anova(lm(TROC~Agemonths,data=Loss))[1,5])`; Fig. 1B)

Both early-life telomere length and rate of loss varied significantly among breeding seasons (one-way ANOVA, telomere length: F = `r round2(anova(lm(RTL~factor(FieldPeriodID),data = juv))[1,4])`; P `r round3(anova(lm(RTL~factor(FieldPeriodID),data = juv))[1,5])`; Fig, 1C; telomere loss: F = `r round2(anova(lm(TROC~factor(FieldPeriodID),data = Loss))[1,4])`; P = `r round3(anova(lm(TROC~factor(FieldPeriodID),data = Loss))[1,5])`; Fig, 2). Variation in median telomere length over breeding seasons in fledglings and subadults was not related temporal variation in  territory quality (linear regression, R^2^ = `r round2(summary(lm(RTL~TQ,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(RTL~TQ,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(RTL~TQ,data = juvseason,na.action=na.exclude))[1,5])`), island-wide food availability (R^2^ = `r round2(summary(lm(RTL~Insect,data=subset(juvseason,Insect<9),na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(RTL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(RTL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,5])`) or population density (R^2^ = `r round2(summary(lm(RTL~Density,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(RTL~Density,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(RTL~Density,data = juvseason,na.action=na.exclude))[1,5])`).

#####Early-life environment and early-life telomere dynamics
The top model explaining within-season variation in early-life RTL contained age, territory quality, the number of helpers in the natal territory, tarsus length, and interactions between age and the number of helpers and tarsus length (Table S1). All the the terms in the top model were significant (FIg. 2A), and the top model was much better supported than a null model ($\Delta$AICc = `r round2(AICc(lmer(RTL~1+(1|LayYear)+(1|TerritoryID),data=juv,REML = FALSE))-AICc(lmer(RTL~Agemonths+Helper+TQ+Tarsus+Agemonths*Helper+Agemonths*Tarsus+(1|LayYear)+(1|TerritoryID),data=juv,REML = FALSE)))`). Tarsus length was negatively related to RTL in nestlings, unrelated to RTL in fledglings, and positively related to RTL in subadults (Fig. 2B). The number of helpers was positvely associated with RTL in nestlings, but not in fledglings or subadults (Fig. 2C). Finally, territory quality was negatively associated with RTL, and this effect was most pronounced in fledglings and subadults (Fig. 2D).

With the longitudinal data, the top model explaining telomere loss contained age, sex and body mass (Table S2). This model was a significantly better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(TROC~1+(1|LayYear),data=subset(Loss,TimeDiff>0.3),REML = FALSE))-AICc(lmer(TROC~Agemonths+Sex+BodyMass+(1|LayYear),data=subset(Loss,TimeDiff>0.3),REML = FALSE)))`), but was not a significantly better fit than a model containing only age ($\Delta$AICc = `r round2(AICc(lmer(TROC~Agemonths+(1|LayYear),data=subset(Loss,TimeDiff>0.3),REML = FALSE))-AICc(lmer(TROC~Agemonths+Sex+BodyMass+(1|LayYear),data=subset(Loss,TimeDiff>0.3),REML = FALSE)))`). Furthermore, the sex and body mass effects both had confidence intervals overlapping zero (Fig. 3A), with very little signal of a pattern in the raw data (Fig. 3B, 3C).

@@

Telomere length in early life did not affect survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~RTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~RTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and there was no interaction effect between RTL and juvenile age on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~RTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~RTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`). To separate out cohort-level RTL affected survival we calculated cohort-level RTL (i.e. the mean for each cohort) and added this term to a logistic regression. However, we found no effect of RTL on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and no interaction with juvenile age (estimate = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`)

Cox regression also showed no effect of individual-level or cohort-level RTL on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL+Fledged)))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL)))$coefficients[1,5])`;  cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,6])`).

@@

###Discussion
Here we use the long-term study of a closed population of Seychelles warblers to assess the relationships between early-life conditions, telomere length and survival. We find that while RTL varies

The clearest result from our study is that RTL varies among cohorts. Very few studies have shown that temporal variation in telomere dynamics occurs in natural populations, and to our knowledge the studies that have done so were limited to just two seasons [@Mizutani2013; @Watson2015]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics does occur at the population level over longer time periods.

@@

In the Seychelles warbler we found only weak evidence that any of the social or environmental vairables we measured affect juvenile RTL. This is surprising a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival. Possible explanations for this finding include i) temporal variation in RTL in our data is so strong that we are unable to detect spatial trends, ii) within-cohort variation in RTL is explained by a variable that we have not included in our analyses, or iii) that effects are generally weak and levels of noise in our telomere measurments and/or ecological data preclude detection of significant effects. Future research should therefore examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence in natural populations [@Becker2015; @Asghar2014].

Our data suggest that, in addition to age class, tarsus length and season had weak effects on juvenile RTL (Figs 3, 4). The tarsus effect most likely reflects the fact that in passerine birds tarsus length is correlated with age during the nestling stage [@Ricklefs1976]. Telomere loss is most rapid early in life due to ongoing cell replication [@Frenck1998], and a negative correlation between RTL and body size in early life is therefore expected. Indeed, that age class affects both tarsus length and RTL can be clearly seen in our data (Fig. 4B). The difference in RTL between birds born in summer and winter seasons is more surprising, especially given the directionality of the effect. We find that birds born in winter seasons having longer telomeres (Fig. 4A), while the opp. One possible explanation for this is that parents in good condition are more likely to breed

### 2016-02-29 Manuscript/SW_early_life_telomeres_SOM.Rmd
More major edits

#Run the model
lm1 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|LayYear)+(1|PlateID)+(1|BirdID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

```{r Table S2}
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)

#Model averaging
t1 <- as.matrix(subset(lm3, delta<=6))

@@

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:7)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

**Figure S1**
```{r Figure S1}
#####################
#4A
#####################
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.LogTL'] <- 'Initial TL'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

>>>>

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|LayYear)+(1|TerritoryID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

```{r Table S2}
temp <- subset(Loss,TimeDiff>0.3)

#Run the model
lm1 <- lmer(TROC ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + RTL + (1|LayYear),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

nams[nams == 'z.RTL'] <- 'RTL'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:15)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

@@

**Figure S1**
```{r Figure S1}
#####################
#4A
#####################
temp <- subset(Loss,TimeDiff>0.3)
temp <- subset(temp,Agemonths == 1)
temp <- subset(temp,TROC<2)
temp <- subset(temp,TQI<4)

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.RTL'] <- 'Initial TL'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

ggplot(temp,aes(x = RTL,y = TROC,col=factor(Helper)))+
geom_point()+
stat_smooth(method = 'lm')
